On Math

### 2025-05-06 cards/paper.tex
law of cosines
more updates:
- graphing cube functions
- normal line functions
- tangent line functions
- chain rule for differentiation

@@

\subsection{Exponential functions}
\subsection{Logarithmic functions}
\subsection{Cubic curves functions}
general function
\begin{equation}
  f(x) = ax^3 + bx^2 + cx + d
\end{equation}
range:
\begin{equation}
  y
\end{equation}

@@

\section{Polynomials}
\subsection{Polynomial identities}
\begin{flalign*}
  & (a + b)^2 = a^2 + 2ab + b^2 & \\
  & (a - b)^2 = a^2 - 2ab + b^2 & \\
  & (a + b)^3 = a^3 + 3a^2b + 3ab^2 + b^3 & \\
  & (a - b)^3 = a^3 - 3a^2b + 3ab^2 - b^3 & \\
  & a^2 - b^2 = (a + b)(a - b) & \\
  & a^3 - b^3 = (a - b)(a^2 + ab + b^2) & \\
  & a^3 + b^3 = (a + b)(a^2 - ab + b^2) &
\end{flalign*}
\subsection{End behavior of general polynomials}
Given polynomial
\begin{equation}
  f(x) = a_nx^n + a_{n-1}x^{n-1} + ... +a_1x + a_0
\end{equation}
with leading coefficient $a \neq 0$ the end behavior is always
\begin{equation}
  f(x) \rightarrow \infty
\end{equation}
or
\begin{equation}
  f(x) \rightarrow -\infty
\end{equation}
as
\begin{equation}
  x \rightarrow \pm \infty
\end{equation}
If the leading coefficient is positive then
$f(x) \rightarrow \infty$ as $x \rightarrow \infty$,
if negative then
$f(x) \rightarrow -\infty$ as $x \rightarrow \infty$
If leading exponent is \textbf{even}, then the end behavior as $x \rightarrow \infty$ \textbf{is the same} as the end behavior as $x \rightarrow -\infty$.
Otherwise, if the leading exponent is \textbf{odd}, then the end behavior as $x \rightarrow +\infty$ is the opposite as $x \rightarrow -\infty$.
\begin{table}[htbp]
  \centering
  \begin{tabular}{|l|l|l|l|}
    \hline
    \textbf{$n$} & \textbf{$a_n$} & \textbf{$x \rightarrow -\infty$} & \textbf{$x \rightarrow +\infty$} \\
    \hline
    even & $>0$ & $f(x) \rightarrow +\infty$ & $f(x) \rightarrow +\infty$ \\
    \hline
    even & $<0$ & $f(x) \rightarrow -\infty$ & $f(x) \rightarrow -\infty$ \\
    \hline
    odd & $>0$ & $f(x) \rightarrow -\infty$ & $f(x) \rightarrow +\infty$ \\
    \hline
    odd & $<0$ & $f(x) \rightarrow +\infty$ & $f(x) \rightarrow -\infty$ \\
    \hline
  \end{tabular}
  \caption{End behavior of polynomials}
  \label{tab:end_behavior_of_polynomials}
\end{table}
Mnemonic trick:
substitute $+1$ and $-1$ into the leading term of polynomial, for example:
\begin{equation}
  \begin{gathered}
    P(x) = 2x^3 - 6x^2 + 8 \\
    \\
    2(+1)^3 = +2 \\
    P(x) \rightarrow +\infty \\
    \text{as} \\
    x \rightarrow +\infty \\
    \\
    2(-1)^3 = -2 \\
    P(x) \rightarrow -\infty \\
    \text{as} \\
    x \rightarrow -\infty \\
  \end{gathered}
\end{equation}
In general whenever $n$ is an \textit{even natural number} we have
\begin{equation}
  \begin{gathered}
    \lim_{x \to +\infty}x^n = +\infty \\
    \text{and} \\
    \lim_{x \to -\infty}x^n = +\infty
  \end{gathered}
\end{equation}
for $n = 2, 4, 6, ...$ but for \textit{odd natural number} though:
\begin{equation}
  \begin{gathered}
    \lim_{x \to +\infty}x^n = +\infty \\
    \text{and} \\
    \lim_{x \to -\infty}x^n = -\infty
  \end{gathered}
\end{equation}
for $n = 1, 3, 5, ...$. To compute the limit at infinity of general polynomial, we only need
to consider the term with the largest expontent, called the \textbf{leading term}.
All of the other terms can be disregarded.
\begin{equation}
  \begin{gathered}
    \lim_{x \to \infty}(2x^3 - 7x^2 + 6x + 12) = \\
    = \lim_{x \to \infty}(2x^3) = \\
    = \lim_{x \to \infty}(x^3) = \\
    = \infty
  \end{gathered}
\end{equation}
\section{Derivatives}
Average rate of change of the function $f(x)$ on the interval $[a, a+h]$:

\begin{equation}
  \frac{\Delta y}{\Delta x} = \frac{f(a+h) - f(a)}{h}
\end{equation}

Instanteous rate of change of the function at the point $x=a$ is the limit of the average
rate of change as $h \rightarrow 0$. It's denoted by $f'(a)$ and given by:

\begin{equation}
  f'(a) = \lim_{h \rightarrow 0}\frac{f(a+h) - f(a)}{h}
\end{equation}

and its value is equal to the slope of the tangent line to the curve $y = f(x)$ at $x=a$.
Also it can be defined as:
\begin{equation}
  f'(a) = \lim_{x \rightarrow a}\frac{f(x) - f(a)}{x-a}
\end{equation}

\subsection{Differentiation of an exponent function}
\begin{equation}
  \frac{d}{dx} (a^x) = a^x\ln a
\end{equation}
with special case:
\begin{equation}
  \frac{d}{dx} (e^x) = e^x\ln e = e^x * 1 = e^x
\end{equation}
\subsection{Chain rule for exponent function}
\begin{equation}
  \begin{gathered}
    y = e^{3x} \\
    y = e^u \\
    u = 3x \\
    \frac{dy}{dx} = \frac{dy}{du} * \frac{du}{dx} \\
    \frac{d}{du} e^u = e^u \\
    \frac{d}{dx} 3x = 3 \\
    \frac{dy}{dx} = e^u * 3 = 3e^u = 3e^{3x}
  \end{gathered}
\end{equation}

>>>>

\subsection{Law of cosines}
% TODO: image needed
\begin{equation}
  \begin{gathered}
    a^2 = b^2 + c^2 - 2bc\cos{A} \\
    b^2 = a^2 + c^2 - 2ac\cos{B} \\
    c^2 = a^2 + b^2 - 2ab\cos{C}
  \end{gathered}
\end{equation}

@@

\subsection{Exponential functions}
\subsection{Logarithmic functions}
\subsection{Cubic curves functions}
general function
\begin{equation}
  f(x) = ax^3 + bx^2 + cx + d
\end{equation}
range:
\begin{equation}
  y \in (-\infty, \infty)
\end{equation}
At least one real root and at most three distinct roots. For example:
\begin{equation}
  \begin{gathered}
    y = x^3 - x \\
    0 = x(x^2 - 1) \\
    0 = x(x - 1)(x + 1) \\
    x = -1, 0, 1
  \end{gathered}
\end{equation}
%TODO image

- for single roots, the curve passes through the $x-axis$.

- touches the $x-axis$ and turns around at the double root.

- for root of multiplicity 3, the graph becomes \textit{tangent} (lies parallel) to the $x-axis$ and crosses it.

@@

\section{Polynomials}
\subsection{Polynomial identities}
\begin{flalign*}
  & (a + b)^2 = a^2 + 2ab + b^2 & \\
  & (a - b)^2 = a^2 - 2ab + b^2 & \\
  & (a + b)^3 = a^3 + 3a^2b + 3ab^2 + b^3 & \\
  & (a - b)^3 = a^3 - 3a^2b + 3ab^2 - b^3 & \\
  & a^2 - b^2 = (a + b)(a - b) & \\
  & a^3 - b^3 = (a - b)(a^2 + ab + b^2) & \\
  & a^3 + b^3 = (a + b)(a^2 - ab + b^2) &
\end{flalign*}
\subsection{End behavior of general polynomials}
Given polynomial
\begin{equation}
  f(x) = a_nx^n + a_{n-1}x^{n-1} + ... +a_1x + a_0
\end{equation}
with leading coefficient $a \neq 0$ the end behavior is always
\begin{equation}
  f(x) \rightarrow \infty
\end{equation}
or
\begin{equation}
  f(x) \rightarrow -\infty
\end{equation}
as
\begin{equation}
  x \rightarrow \pm \infty
\end{equation}
If the leading coefficient is positive then
$f(x) \rightarrow \infty$ as $x \rightarrow \infty$,
if negative then
$f(x) \rightarrow -\infty$ as $x \rightarrow \infty$
If leading exponent is \textbf{even}, then the end behavior as $x \rightarrow \infty$ \textbf{is the same} as the end behavior as $x \rightarrow -\infty$.
Otherwise, if the leading exponent is \textbf{odd}, then the end behavior as $x \rightarrow +\infty$ is the opposite as $x \rightarrow -\infty$.
\begin{table}[htbp]
  \centering
  \begin{tabular}{|l|l|l|l|}
    \hline
    \textbf{$n$} & \textbf{$a_n$} & \textbf{$x \rightarrow -\infty$} & \textbf{$x \rightarrow +\infty$} \\
    \hline
    even & $>0$ & $f(x) \rightarrow +\infty$ & $f(x) \rightarrow +\infty$ \\
    \hline
    even & $<0$ & $f(x) \rightarrow -\infty$ & $f(x) \rightarrow -\infty$ \\
    \hline
    odd & $>0$ & $f(x) \rightarrow -\infty$ & $f(x) \rightarrow +\infty$ \\
    \hline
    odd & $<0$ & $f(x) \rightarrow +\infty$ & $f(x) \rightarrow -\infty$ \\
    \hline
  \end{tabular}
  \caption{End behavior of polynomials}
  \label{tab:end_behavior_of_polynomials}
\end{table}
Mnemonic trick:
substitute $+1$ and $-1$ into the leading term of polynomial, for example:
\begin{equation}
  \begin{gathered}
    P(x) = 2x^3 - 6x^2 + 8 \\
    \\
    2(+1)^3 = +2 \\
    P(x) \rightarrow +\infty \\
    \text{as} \\
    x \rightarrow +\infty \\
    \\
    2(-1)^3 = -2 \\
    P(x) \rightarrow -\infty \\
    \text{as} \\
    x \rightarrow -\infty \\
  \end{gathered}
\end{equation}
In general whenever $n$ is an \textit{even natural number} we have
\begin{equation}
  \begin{gathered}
    \lim_{x \to +\infty}x^n = +\infty \\
    \text{and} \\
    \lim_{x \to -\infty}x^n = +\infty
  \end{gathered}
\end{equation}
for $n = 2, 4, 6, ...$ but for \textit{odd natural number} though:
\begin{equation}
  \begin{gathered}
    \lim_{x \to +\infty}x^n = +\infty \\
    \text{and} \\
    \lim_{x \to -\infty}x^n = -\infty
  \end{gathered}
\end{equation}
for $n = 1, 3, 5, ...$. To compute the limit at infinity of general polynomial, we only need
to consider the term with the largest expontent, called the \textbf{leading term}.
All of the other terms can be disregarded.
\begin{equation}
  \begin{gathered}
    \lim_{x \to \infty}(2x^3 - 7x^2 + 6x + 12) = \\
    = \lim_{x \to \infty}(2x^3) = \\
    = \lim_{x \to \infty}(x^3) = \\
    = \infty
  \end{gathered}
\end{equation}
\section{Derivatives}
Average rate of change of the function $f(x)$ on the interval $[a, a+h]$:
\begin{equation}
  \frac{\Delta y}{\Delta x} = \frac{f(a+h) - f(a)}{h}
\end{equation}
Instanteous rate of change of the function at the point $x=a$ is the limit of the average
rate of change as $h \rightarrow 0$. It's denoted by $f'(a)$ and given by:
\begin{equation}
  f'(a) = \lim_{h \rightarrow 0}\frac{f(a+h) - f(a)}{h}
\end{equation}
and its value is equal to the slope of the tangent line to the curve $y = f(x)$ at $x=a$.
Also it can be defined as:
\begin{equation}
  f'(a) = \lim_{x \rightarrow a}\frac{f(x) - f(a)}{x-a}
\end{equation}

\subsection{Tangent line equation}
% TODO image
Equation of the tangent line to the curve $y=f(x)$ at the point $P(x_1, y_1)$ is given by
\begin{equation}
  \begin{gathered}
    y - y_1 = m(x - x_1) \\
    \text{where:} \\
    m = f'(x_1)
  \end{gathered}
\end{equation}

\subsection{Normal line equation}
% TODO image
Equation of the normal line to the curve $y=f(x)$ at the point $P(x_1, y_1)$ is given by
\begin{equation}
  \begin{gathered}
    y - y_1 = m(x - x_1) \\
    \text{where:} \\
    m = -\frac{1}{f'(x_1)}
  \end{gathered}
\end{equation}
Normal line is just negative reciprocal of the slope of the \textbf{tangent line}.

\subsection{Chain rule for differentiation}
Derivative of a composite function $f(g(x))$ is given by:
\begin{equation}
  \frac{df}{dx} = \frac{df}{dg} * \frac{dg}{dx} = \frac{df}{dx}
\end{equation}
Example:
\begin{equation}
  \begin{gathered}
    f(x) = (3x^4 + x^3)^5 \\
    \frac{df}{dx} = ? \\
    f = g^5 \\
    g = 3x^4 + x^3 \\
    \frac{df}{dx} = \frac{df}{dg} * \frac{dg}{dx} \\
    \frac{df}{dg} = 5g^4 \\
    \frac{dg}{dx} = 12x^3 + 3x^2 \\
    \frac{df}{dx} = 5g^4 * (12x^3 + 3x^2) = 5(3x^4 + x^3)^4(12x^3 + 3x^2)
  \end{gathered}
\end{equation}
The chain rule also can be expressed using primes instead of Leibniz notation
\begin{equation}
  f(g(x))' = f'(g(x)) * g'(x)
\end{equation}
\subsection{Differentiation of an exponent function}
\begin{equation}
  \frac{d}{dx} (a^x) = a^x\ln a
\end{equation}
with special case:
\begin{equation}
  \frac{d}{dx} (e^x) = e^x\ln e = e^x * 1 = e^x
\end{equation}
\subsection{Chain rule for exponent function}
\begin{equation}
  \begin{gathered}
    y = e^{3x} \\
    y = e^u \\
    u = 3x \\
    \frac{dy}{dx} = \frac{dy}{du} * \frac{du}{dx} \\
    \frac{d}{du} e^u = e^u \\
    \frac{d}{dx} 3x = 3 \\
    \frac{dy}{dx} = e^u * 3 = 3e^u = 3e^{3x}
  \end{gathered}
\end{equation}
